### 2014-03-04 diploma/kapitel/Einf.tex
add mesh heineb 1-5k

\chapter{Einführung}

History: \cite{whitney} (1957)