### 2014-06-24 diploma/bilder/tikz/ableitungDual1Form.tex
...

\begin{tikzpicture}[>=latex]
  % Coords
  \coordinate (V0) at (2,0);
  \coordinate (V1) at (2,2);
  \coordinate (V2) at (0,1);
  \coordinate (V3) at (4,1);
  \coordinate (V4) at (4,-1);
  \coordinate (V5) at (0,-1);
  % Arrows\tilde{\sigma}
  \draw[line width=1pt, ->, style=dotted]
    (V0) -- (V1);
  \draw[line width=1pt, style=dotted]
    (V1) --  (V2);
  \draw[line width=1pt, ->, style=dotted]
    (V0) -- (V2);
  \draw[line width=1pt, style=dotted]
    (V1) --  (V3);
  \draw[line width=1pt, ->, style=dotted]
    (V0) -- (V3);
   \draw[line width=1pt, ->, style=dotted]
    (V0) -- (V4);
  \draw[line width=1pt, ->, style=dotted]
    (V0) -- (V5);
  \draw[line width=1pt, style=dotted]
    (V5) --  (V2);
  \draw[line width=1pt, style=dotted]
    (V3) --  (V4);
  % Points
  \fill (V0) node[below] {\(v\)} circle (2pt);
  \fill (V1) circle (2pt);
  \fill (V2) circle (2pt);
  \fill (V3) circle (2pt);
  \fill (V4) circle (2pt);
  \fill (V5) circle (2pt);
  \draw[line width=1pt, style=loosely dotted]
     (V5) to[bend right] (V4);
  %circumcenter
  \coordinate (CC1) at (1.333,1);
  \coordinate (CC0) at (2.666,1);
  \coordinate (CC2) at (0.666,0);
  \coordinate (CCm) at (3.333,0);
  \coordinate (CCC0) at (2,1);
  \coordinate (CCC1) at (1,0.5);
  \coordinate (CCC2) at (3,0.5);
  \coordinate (CCC3) at (1,-0.5);
  \coordinate (CCC4) at (3,-0.5);
  \fill (CC0) circle (2pt);
  \fill (CC1) circle (2pt);
  \fill (CC2) circle (2pt);
  \fill (CCm) circle (2pt);
  \fill (CCC0) circle (2pt);
  \fill (CCC1) circle (2pt);
  \fill (CCC2) circle (2pt);
  \fill (CCC3) circle (2pt);
  \fill (CCC4) circle (2pt);
  \draw[line width=1pt, ->] (CC0) -- node[above right] {\(\star\sigma_0^1\)} (CC1);
  \draw[line width=1pt, ->] (CC1) -- node[above left] {\(\star\sigma_1^1\hspace{-6pt}\)}(CC2);
  \draw[line width=1pt, ->] (CCm) -- node[right] {\(\star\sigma_m^1\)} (CC0);
  \draw[line width=1pt,] (CC2) -- node[left] {\(\star\sigma_2^1\)} (CCC3);
  \draw[line width=1pt,->] (CCC4) -- node[right] {\(\star\sigma_{m-1}^1\)} (CCm);
  \draw[line width=1pt, style=loosely dotted]
     (CCC3) to[bend right] (CCC4);

\useasboundingbox ([shift={(1mm,1mm)}]current bounding box.north east) rectangle ([shift={(-1mm,-1mm)}]current bounding box.south west);
\end{tikzpicture}