### 2014-08-26 diploma/kapitel/Einf.tex
...

\pagenumbering{arabic}
\setcounter{page}{1} 
\setcounter{chapter}{-1}
\chapter{Einführung}

History: \cite{whitney} (1957)