### 2014-09-21 verteidigung/bilder/tikz/Kette1.tex
...

\begin{tikzpicture}[>=latex, line width=1pt, scale=0.9]
  % Coords
  \coordinate (V0) at (2,0);
  \coordinate (V1) at (2,2);
  \coordinate (V2) at (0,1);
  \coordinate (V3) at (4,1);
  \coordinate (V4) at (4,-1);
  \coordinate (V5) at (0,-1);
 \coordinate (V6) at (2,-2);

\fill[opacity=0.4]  (V1) -- (V3) -- (V4) -- (V6) -- (V5) -- (V2);

  % Arrows\tilde{\sigma}
  \draw[red,->](V1) --node[left, black]{\(\sigma^1_0\)} node[right, black]{\small 1} (V0);
  \draw(V1) --  (V2);
  \draw(V0) -- (V2);
  \draw(V1) --  (V3);
  \draw(V0) -- (V3);
   \draw(V0) -- (V4);
  \draw[red,->](V0) --node[above left=-3pt, black]{\(\sigma^1_1\)} node[below right=-3pt, black]{\small -1} (V5);
  \draw(V5) --  (V2);
  \draw(V3) --  (V4);
 \draw(V4) -- (V6);
\draw(V0) -- (V6);
\draw[red,->](V5)  --node[below left=-3pt, black]{\(\sigma^1_2\)} node[above right=-3pt, black]{\small 2}(V6);
  % Points
  \fill (V0) circle (2pt);
  \fill (V1) circle (2pt);
  \fill (V2) circle (2pt);
  \fill (V3) circle (2pt);
  \fill (V4) circle (2pt);
  \fill (V5)circle (2pt);
 \fill (V6)circle (2pt);

\end{tikzpicture}