### 2015-09-11 stuff/dirfields/DirFields.tex
...

\documentclass{scrartcl}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{dsfont}

\newcommand{\exd}{\mathbf{d}}
\newcommand{\excod}{\exd^{*}} %or \delta
\newcommand{\Div}{\text{Div}}
\newcommand{\Rot}{\text{Rot}}

\newcommand{\R}{\mathds{R}}

% IDENTIFIERS for math mode:

%surface (manifold) -> M, S, \Omega 
\newcommand{\M}{M}
%volume element (volume measure) -> \mu, dA, d\Omega
\newcommand{\dA}{dA}
%director field (contra-vector)
\newcommand{\p}{\mathbf{p}}
%director field (co-vector, flat of contra-vector)
\newcommand{\pfl}{\mathbf{p}^{\flat}}

\begin{document}

\section{Frank Oseen Energy}

In \( \R^{3} \):
\begin{align}\label{eqFO1}
  E = \frac{1}{2}\int_{\Omega} K_{1}\left( \nabla\cdot\p \right)^{2} 
                        + K_{2}\left( \p\cdot\left[ \nabla\times\p \right] \right)^{2}
                        + K_{3}\left\| \p\times\left[ \nabla\times\p \right] \right\|^{2} dV
\end{align}
With the Langrange identity for the \( K_{3} \)-term, we cann rewrite \eqref{eqFO1} to
\begin{align}\label{eqFO2}
  E = \frac{1}{2}\int_{\Omega} K_{1}\left( \nabla\cdot\p \right)^{2} 
                        + (K_{2}-K_{3})\left( \p\cdot\left[ \nabla\times\p \right] \right)^{2}
                        + K_{3}\left\| \p \right\|^{2}\left\| \nabla\times\p \right\|^{2} dV
\end{align}
If we restrict \eqref{eqFO2} to a 2-dimensional Manifold \( \M\subset\Omega \) 
and postulate that \( \p\in T_{X}\M \) is a normalized tangential vector in \( X\in\M \), we get
\begin{align}
  E = \frac{1}{2} \int_{\M} K_{1}\left( \Div\p \right)^{2} + K_{3}\left( \Rot \p \right)^{2} \dA 
\end{align}
In terms of exterior calculus with the corresponding 1-form \( \pfl\in\Lambda^{1}(M) \), we obtain
\begin{align}
  E = \frac{1}{2} \int_{\M} K_{1}\left( \excod\p \right)^{2} + K_{3}\left( \exd \p \right)^{2} \dA
\end{align}
where \( \excod := -* \exd * \) is the \( L^{2} \)-orthogonal operator of the exterior derivative \( \exd \).

\end{document}