### 2015-09-16 stuff/dirfields/DirFields.tex
...

\documentclass{scrartcl}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{dsfont}

\usepackage{hyperref}

\newcommand{\exd}{\mathbf{d}}
\newcommand{\excod}{\exd^{*}} %or \delta
\newcommand{\Div}{\text{Div}}
\newcommand{\Rot}{\text{Rot}}

\newcommand{\R}{\mathds{R}}

% IDENTIFIERS for math mode:

%surface (manifold) -> M, S, \Omega 
\newcommand{\M}{M}
%volume element (volume measure) -> \mu, dA, d\Omega
\newcommand{\dA}{dA}
%director field (contra-vector)
\newcommand{\p}{\mathbf{p}}
\newcommand{\q}{\mathbf{q}}
%director field (co-vector, flat of contra-vector)
\newcommand{\pfl}{\mathbf{p}^{\flat}}
\newcommand{\qfl}{\mathbf{q}^{\flat}}
%discrete director field (co-vector, flat of contra-vector)
\newcommand{\pflh}{\mathbf{p}^{\flat}_{h}}
%discrete director field (contra-vector)
\newcommand{\ph}{\mathbf{p}_{h}}
%discrete PD director field (co-vector, flat of contra-vector)
\newcommand{\PDpflh}{\underline{\mathbf{p}}^{\flat}_{h}}
%discrete director field (contra-vector)
\newcommand{\PDph}{\underline{\mathbf{p}}_{h}}
%Frank Oseen Energy (without Lagrange term for normalizing)
\newcommand{\EOS}{E_{\text{OS}}}
%Normalizing energy
\newcommand{\EN}{E_{n}}
%Laplace-Beltrami or Rot-Rot-Laplace
\newcommand{\LB}{\Delta^{\text{\tiny RR}}}
%Laplace-CoBeltrami or Grad-Div-Laplace
\newcommand{\LCB}{\Delta^{\text{\tiny GD}}}
%Laplace-deRham
\newcommand{\LDR}{\Delta^{\text{\tiny dR}}}
%discrete Laplace-Beltrami or Rot-Rot-Laplace
\newcommand{\LBh}{\LB_{h}}
%Laplace-CoBeltrami or Grad-Div-Laplace
\newcommand{\LCBh}{\LCB_{h}}
%Laplace-deRham
\newcommand{\LDRh}{\LDR_{h}}
%Landau symbol
\renewcommand{\O}{\mathcal{O}}
%Vertices
\newcommand{\V}{\mathcal{V}}
%Edges
\newcommand{\E}{\mathcal{E}}
%Faces
\newcommand{\F}{\mathcal{F}}
%edge vector
\newcommand{\e}{\mathbf{e}}
%vertex vector
\renewcommand{\v}{\mathbf{v}}

\begin{document}

\tableofcontents

\section{Frank Oseen energy}

In \( \R^{3} \):
\begin{align}\label{eqFO1}
  \EOS = \frac{1}{2}\int_{\Omega} K_{1}\left( \nabla\cdot\p \right)^{2} 
                        + K_{2}\left( \p\cdot\left[ \nabla\times\p \right] \right)^{2}
                        + K_{3}\left\| \p\times\left[ \nabla\times\p \right] \right\|^{2} dV
\end{align}
With the Langrange identity for the \( K_{3} \)-term, we cann rewrite \eqref{eqFO1} to
\begin{align}\label{eqFO2}
  \EOS = \frac{1}{2}\int_{\Omega} K_{1}\left( \nabla\cdot\p \right)^{2} 
                        + (K_{2}-K_{3})\left( \p\cdot\left[ \nabla\times\p \right] \right)^{2}
                        + K_{3}\left\| \p \right\|^{2}\left\| \nabla\times\p \right\|^{2} dV
\end{align}
If we restrict \eqref{eqFO2} to a 2-dimensional Manifold \( \M\subset\Omega \) 
and postulate that \( \p\in T_{X}\M \) is a normalized tangential vector in \( X\in\M \), we get
\begin{align}
  \EOS = \frac{1}{2} \int_{\M} K_{1}\left( \Div\p \right)^{2} + K_{3}\left( \Rot \p \right)^{2} \dA 
\end{align}
In terms of exterior calculus with the corresponding 1-form \( \pfl\in\Lambda^{1}(M) \), 
,i.e. \( \left( \pfl \right)^{\sharp} = \p\), we obtain
\begin{align}
  \EOS = \frac{1}{2} \int_{\M} K_{1}\left( \excod\pfl \right)^{2} + K_{3}\left( *\exd \pfl \right)^{2} \dA
\end{align}
where the exterior coderivative \( \excod := -* \exd * \) is the \( L^{2} \)-orthogonal operator of the exterior derivative \( \exd \). 
(Note \( \Div\p = -\excod\pfl \) and \( \Rot\p = *\exd\pfl \))

  \subsection{Functional derivative}

  With the  \( L^{2} \)-orthogonality of the exterior derivative and coderivative and a arbitrary \( \alpha\in\Lambda^{1}(M) \) we get
  \begin{align}
    \int_{\M} \left\langle\frac{\delta\EOS}{\delta\pfl} , \alpha \right\rangle \dA
      &= \lim_{\epsilon\rightarrow 0} \frac{1}{\epsilon}\left( \EOS\left[ \pfl + \epsilon\alpha \right] - \EOS\left[ \pfl\right]\right) \\
      &=  \lim_{\epsilon\rightarrow 0} \frac{1}{2\epsilon} \int_{\M} 
                K_{1}\left( 2\epsilon \left( \excod\pfl \right)\left( \excod\alpha \right) 
                            + \epsilon^{2}\left( \excod\alpha \right)^{2} \right) \\
      &\phantom{= \lim_{\epsilon\rightarrow 0} \frac{1}{2\epsilon} \int_{\M}}
                +K_{3}\left(2\epsilon\left\langle \exd\pfl, \exd\alpha \right\rangle
                            + \epsilon^{2}\left\| \exd\alpha \right\|^{2}\right)\dA \\
      &= -\int_{\M} K_{1} \left\langle \LCB\pfl, \alpha \right\rangle + K_{3} \left\langle \LB\pfl, \alpha \right\rangle \dA \\
      &= \int_{\M} \left\langle -\left( K_{1}\LCB + K_{3}\LB\right)\pfl , \alpha \right\rangle \dA
  \end{align}
  where \( \LB = -\excod\exd = *\exd *\exd  \) is the Vector-Laplace-Beltrami-Operator or Rot-Rot-Laplace 
  and \( \LCB = - \exd\excod = \exd * \exd * \) is the Vector-Laplace-CoBeltrami-Operator or Grad-Div-Laplace.
  Hence, for a One-Constant-Approximation \( K_{1} = K_{3} =: K_{0} \), we obtain
  \begin{align}
    \int_{\M} \left\langle\frac{\delta\EOS}{\delta\pfl} , \alpha \right\rangle \dA
     &= \int_{\M} \left\langle K_{0}\LDR\pfl , \alpha \right\rangle \dA
  \end{align}
  where \( \LDR = -\LB - \LCB = \excod\exd + \exd\excod\) is the Laplace-de Rham operator.

  \subsection{Unit vector invariance}

  If \( \p\in T_{X}\M \) is a unit vector on \( \M \), we can describe all unit vectors \(  \) in \( X\in\M \) as a rotation in the tangential space
  with angle \( \phi\in\R \):
  \begin{align}
    \q = \cos\phi\p + \sin\phi\left( *\p \right)
  \end{align}
  \( *\p = \left( *\pfl \right)^{\sharp}\) is the Hodge dual of \( \p \), i.e. a quarter rotation of \( \p \). 
  For a space independent angle \( \phi \), i.e. \( \exd\phi=0 \), straight forward calculations implies
  \begin{align}
     \left\| \Rot (*\p) \right\| &=  \left\| *\exd*\pfl \right\| = \left\| \Div\p \right\|\\
     \left\| \Div (*\p) \right\| &= \left\| *\exd**\pfl  \right\| = \left\| *\exd\pfl  \right\| = \left\|\Rot\p\right\|\\ 
     \left\| \Rot\q \right\|^{2} &= \left\| *\exd\qfl \right\|^{2} = \left\| \exd\qfl \right\|^{2} \\
                                      &=  \cos^{2}\phi\left\| \Rot\p \right\|^{2} 
                                        + \sin^{2}\phi\left\| \Div\p \right\|^{2}
                                        + 2\cos\phi\sin\phi\left\langle \exd\pfl, \exd * \pfl \right\rangle\\
    \left\| \Div\q \right\|^{2} &= \left\| *\exd*\qfl \right\|^{2} = \left\| \exd*\qfl \right\|^{2}\\
                                     &= \cos^{2}\phi\left\| \Div\p \right\|^{2} 
                                        + \sin^{2}\phi\left\| \Rot\p \right\|^{2}
                                        - 2\cos\phi\sin\phi\left\langle \exd\pfl, \exd * \pfl \right\rangle  
  \end{align}
  Finally, we get for the One-Constant-Approximation of the Frank-Oseen-Energy
  \begin{align}
    \EOS[\q] = \EOS[\p]
  \end{align}

\section{Normalizing energy}

To constrain \( \p \) is normalized, we add
\begin{align}
  \EN = \int_{\M} \frac{K_{n}}{4} \left( \left\| \p \right\|^{2} - 1 \right)^{2} \dA
\end{align}
to the Frank Oseen energy.
Note that the norm defined by the metric \( g \) on the manifold \( \M \) is invariant regarding lowering or rising the indices, i.e.
\begin{align}
  \left\| \p \right\|^{2} = p^{i}g_{ij}p^{j} = p_{i} g^{ij} p_{j} = \left\| \pfl \right\|^{2}
\end{align}

  \subsection{Functional derivative}

  By variating \( \pfl \) under the norm with an arbitrary \( \alpha\in\Lambda^{1}(\M) \), we obtain
  \begin{align}
    \left\| \pfl + \epsilon \alpha \right\|^{2} &= \left\| \pfl \right\|^{2} 
                                                  + 2\epsilon\left\langle \pfl, \alpha \right\rangle
                                                  + \epsilon^{2}\left\| \alpha \right\|^{2}
  \end{align}
  If we are only interesting in linear terms (in \( \epsilon \)), this leads to
  \begin{align}
    \left( \left\| \pfl + \epsilon \alpha \right\|^{2} - 1 \right)^{2}
            &= \left( \left\| \pfl \right\|^{2} - 1  + 2\epsilon\left\langle \pfl, \alpha \right\rangle 
                  + \O\left( \epsilon^{2} \right) \right)^{2}\\
            &= \left( \left\| \pfl \right\|^{2} - 1 \right)^{2} 
                + 4\epsilon\left(\left\| \pfl \right\|^{2} - 1 \right)\left\langle \pfl, \alpha \right\rangle
                + \O\left( \epsilon^{2} \right)
  \end{align}
  Hence, we get for the functional derivative of \( \EN \)
  \begin{align}
    \int_{\M} \left\langle\frac{\delta\EN}{\delta\pfl} , \alpha \right\rangle \dA
      &= \lim_{\epsilon\rightarrow 0} \frac{1}{\epsilon}\left( \EN\left[ \pfl + \epsilon\alpha \right] - \EN\left[ \pfl\right]\right) \\
      &= \int_{\M} \left\langle K_{n} \left( \left\| \pfl \right\|^{2} - 1 \right)\pfl , \alpha \right\rangle \dA
  \end{align}

\section{Model equations}
  To minimize the energy \( E := \EOS + \EN \) we choose a time evolving approach. 
  Hence, with the fundamental lemma of calculus of variations, we will use the time depended differential equation in terms of exterior
  calculus
  \begin{align}
    \partial_{t}\pfl = -\frac{\delta E}{\delta\pfl}
                     = -K_{0}\LDR\pfl - K_{n} \left( \left\| \pfl \right\|^{2} - 1 \right)\pfl
  \end{align}
  or in general, if we don't want to use the One-Constant-Approximation,
  \begin{align}
    \partial_{t}\pfl = \left( K_{1}\LCB + K_{3}\LB\right)\pfl - K_{n} \left( \left\| \pfl \right\|^{2} - 1 \right)\pfl
  \end{align}
  Note that if we apply the Hodge operator on the whole equations, we get the Hodge dual equations
  \begin{align}
    \partial_{t}(*\pfl) &= -K_{0}\LDR (*\pfl) - K_{n} \left( \left\| \pfl \right\|^{2} - 1 \right)(*\pfl) \\
                        &=  \left( K_{1}\LB + K_{3}\LCB\right) (*\pfl) - K_{n} \left( \left\| \pfl \right\|^{2} - 1 \right)(*\pfl)
  \end{align}
  which are very useful for the DEC discretization later in context.
  But this leads to pay attention, because only in the first line (One-Constant-Approximation) we see, 
  that the Hodge dual equations in \( *\pfl \) is the same as the primal equation in \( \pfl \). 
  In the general case (second line), we must "swap" the Laplace operators.

\section{A DEC approach}

For further information see for example \cite{whitney,hirani}.

  \subsection{Surface Mesh}
    ...wellcentered manifoldlike simplicial complex, bla, bla, blub...

  \subsection{Discrete 1-forms}

  The main concept to represent a discrete 1-form \( \pflh\in\Lambda_{h}(K) \) is to approximate the contraction of the continuous 1-Form 
  \( \pfl\in\Lambda(M) \) on all edges \( e\in\E \)
  \begin{align}
    \pflh(e) := \int_{\pi(e)}\pfl \approx \int_{0}^{1} \pfl_{X_{e}(\tau)}\left( \dot{X}_{e}(t) \right) dt = \pfl_{X_{e}(\tau)}(\e)
  \end{align}
  where \( \pi:K\rightarrow \M \) is the glueing map, who project the elements of the surface mesh to the manifold.
  \( X_{e}(t) = t\v_{2} + (1-t)\v_{1} \) is the linear barycentric parametrisation of the edge \( e=\left[ v_{1}, v_{2} \right]\).
  The existence of a intermediate value \( \tau\in\left[ 0,1 \right] \), so that \( \e\in T_{X_{e}(\tau)}M \),
  is ensured by the mean value theorem.
  Other discrete forms of arbitrary degree and theirs hodge duals can be interpreted in a similarly way.

  \subsection{Discrete Laplace operators}

    In the discrete exterior calculus discrete Operators are defined by successively interpretation of the basic operations on the forms, 
    like the Hodge operator \( * \) or the exterior derivative \( \exd \), as geometric operators on the simplices, 
    like the Voronoi dual operator \( \star \) or the boundary operator \( \partial \) (see \cite{hirani}).
    This results for example to a discrete definition of \( \LB \) for a discrete 1-form \( \pflh\in\Lambda^{1}_{h}(M) \) on a edge 
    \(e\in\E \)
    \begin{align}
      \LBh\pflh(e) &:=  \left(*\exd*\exd \pflh\right)(e)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \left(\exd * \exd  \pflh\right)(\star e) \\
                    &=  -\frac{\left| e \right|}{\left| \star e \right|} \left(* \exd  \pflh\right)(\partial\star e)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} s_{f,e} \left(* \exd \pflh\right) (\star f) \\
                    &= -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|} 
                                \left(\exd\pflh\right)(f)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|} \pflh(\partial f)\\
                    &= -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|}
                                \sum_{\tilde{e}\prec f} s_{f,\tilde{e}} \pflh(\tilde{e})
    \end{align}
    or for \( \LCB \)
    \begin{align}
      \LCBh\pflh(e) &:= \left( \exd * \exd * \pflh\right) (e)
                      = \left( * \exd *\pflh \right) (\partial e) \\
                     &= \sum_{v\prec e} s_{v,e} \left( * \exd * \pflh \right)(v) 
                      = \sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \left( \exd * \pflh \right)(\star v)\\
                     &= \sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \left( * \pflh \right)(\partial\star v)
                      = -\sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \sum_{\tilde{e}\succ v}
                                s_{v,\tilde{e}} \left( * \pflh \right) (\star \tilde{e}) \\
                     &= -\sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \sum_{\tilde{e}\succ v}
                                s_{v,\tilde{e}} \frac{\left| \star\tilde{e} \right|}{\left| \tilde{e} \right|}\pflh(\tilde{e})
    \end{align}
    where the sign \( s_{f,e} \) is +1 if the face \( f\succ e \) is in the left of the edge \( e \) (-1 otherwise)
    and \( s_{v,e} \) is +1 if the edge \( e\succ v \) points to the vertex \( v \) and -1 if \( e \) points away from \( v \).
    We see \( \LBh \), \( \LCBh \) and also \( \LDRh := -\LBh -\LCBh \) are linear operators in \( \pflh(\tilde{e}) \)
    and therefor results in sparse matrices if the \( \pflh(\tilde{e}) \) are our degree of freedoms.
    ...explainable figures...

  \subsection{Discrete norm}

    Approximating the norm \( \left\| \pfl \right\| \) on a edge \( e\in\E \) is not so easy like the development of discrete linear
    operators. 
    We only know how \( \pflh \) "act" on a single edge, so \( \pflh \) gives us only one dimensional informations.
    In other words, we only know the proportion \( \ph \cdot \e = \pflh (e)\) of the discrete contra vector field \( \ph = \left( \pflh \right)^{\sharp} \) in the
    \( \e \) direction, but we don't know the length of \( \ph \) defined on this edge.

    ...some averaging techniques and why these sucks...

    One way out is to rise the dimension of the discrete 1-forms, therefor we introduce some bases at the intersection 
    \( c(e) = e\cap (\star e) \) of a edge and its dual edge to describe discrete contra- and covariant vector fields in a local (flat)
    coordinate system.

    The basis for contravariant vectors at \( c(e) \) is composed of
    \begin{align}
      \partial_{e}X &:= \e = \v_{2} - \v_{1} \\
      \partial_{\star e}X &:= *\e =  c(f_{2}) - c(f_{1})
    \end{align}
    if \( e=\left[ v_{1}, v_{2} \right] \) and the face \( f_{1} \) lay right and \( f_{2} \) left of \( e \) (...figure...). 
    This definitions are consistent with the canonical basis, if the position \( X \) is a barycentric parametrisation of the edge \( e \)
    resp. its dual, i.e. for
    \begin{align}
      X_{e}(t_{e}) &= t_{e}\v_{2} + \left( 1 - t_{e} \right)\v_{2} \text{ for } t\in[0,1]\\
      X_{\star e}(t_{\star e}) &=\begin{cases}
                        2t_{\star e}c(e) + \left( 1 - 2t_{\star e} \right) c(f_{1}) & \text{if } t_{\star e}\in\left[ 0 , \frac{1}{2} \right] \\
                        \left(1 - t_{\star e}\right) c(e) + \left( 2t_{\star e} - 1 \right) c(f_{1}) 
                                & \text{if } t_{\star e}\in \left[\frac{1}{2} , 1\right]
                      \end{cases}
    \end{align}
    holds \( \partial_{e}X = \partial_{t_{e}}X_{e} \) 
    and \( \partial_{\star e}X = \frac{1}{2} \left( \partial_{t_{\star e}}X_{\star e}|_{\left[ 0 , \frac{1}{2} \right]} 
                                                   + \partial_{t_{\star e}}X_{\star e}|_{\left[\frac{1}{2}, 1 \right]}\right) \).
    Therefore we get the local metric tensor
    \begin{align}
      g_h(e) &= \left| e \right|^{2} \left( dx^{e} \right)^{2} + \left| \star e \right|^{2} \left( dx^{\star e} \right)^{2}
    \end{align}
    where \( \left\{  dx^{e}, dx^{\star e} \right\} \) are the dual base of \( \left\{ \partial_{e}X, \partial_{\star e}X \right\} \),
    i.e. \( dx^{i}\left(\partial_{j}X\right) = \delta^{i}_{j} \) for \( i,j\in\left\{ e, \star e \right\} \).
    This gives us the great possibility to define
    \begin{align}
      \PDpflh(e) &:= \pflh(e)dx^{e} + \pflh(\star e)  dx^{\star e} \\
                   &= \pflh(e)dx^{e} - \frac{\left| \star e \right|}{\left| e \right|} \left(*\pflh\right)(e)  dx^{\star e} \\
      \PDph(e) &:= \left(\PDpflh(e)\right)^{\sharp} 
                 = \frac{1}{\left| e \right|^{2}} \pflh(e)\partial_{e}X + \frac{1}{\left|\star e \right|^{2}} \pflh(\star e)\partial_{\star e}X \\
               & =  \frac{1}{\left| e \right|} \left( \frac{1}{\left| e \right|} \pflh(e)\partial_{e}X 
                                                     -\frac{1}{\left|\star e \right|} \left(*\pflh\right)(e) \partial_{\star e}X\right)
    \end{align}
    We call \( \PDpflh(e) \) the discrete Primal-Dual-1-form (short PD-1-form) and \( \PDph \) the discrete Primal-Dual-vector field
    (short PD-vector field).
    Hence, we get for the square of the norm on \( c(e) \) by contract the PD-1-Form with its corresponding PD-Vector
    \begin{align}
      \left\| \PDpflh \right\|^{2}(e) := \left( \PDpflh\left( \PDph \right) \right)(e)
                                       = \frac{1}{\left| e \right|^{2}}\left( \left[ \pflh(e) \right]^{2} 
                                                                             +\left[ \left(*\pflh\right)(e) \right]^{2}\right)
    \end{align}

\bibliography{bibl}
\bibliographystyle{alpha}

\end{document}