### 2015-09-24 TalkDECPD/bilder/tikz/SCircum.tex
...

\begin{tikzpicture}[>=latex, scale=2, line width=1pt]
          % Coords
          \coordinate (V0) at (0,0);
          \coordinate (V1) at (2,0);
          \coordinate (V2) at (1,2);
          \coordinate (C01) at (1,0);
          \coordinate (C02) at (0.5,1);
          \coordinate (C12) at (1.5,1);
          \coordinate (C) at (1,0.75);
          % Arrows
          \draw[opacity=1]
            (V0) -- (V1);
          \draw[opacity=1]
            (V1) -- (V2);
          \draw[opacity=1]
            (V2) -- (V0);

		%\fill[red] (V0) -- (V1) -- (V2) -- (V0);

          % Points
          \fill[opacity=1](V0) node[below left] {\( v \)};
		  \fill[opacity=1, red](V0) node[above left] {\( c(v) \)} circle (1.5pt);
		  \fill[opacity=0] (V0)  circle (1.5pt);
          \fill[opacity=1] (V1)  circle (1.5pt);
          \fill[opacity=1](V2)   circle (1.5pt);
          \fill[opacity=0] (C02)  circle (1.5pt);
          \fill[opacity=1] (C01) node[below] {\( e \)};
		  \fill[opacity=1, red] (C01) node[above] {\( c(e) \)} circle (1.5pt);
          \fill[opacity=0](C12)  circle (1.5pt);
          \fill[opacity=1](C) node[below] {\( f\)};
		  \fill[opacity=1, red](C) node[above] {\( c(f)\)} circle (1.5pt);

		 \coordinate (CC2) at (1,0.58);
          \draw[opacity=0, color=red, ->] (CC2) + (135:0.2) arc (135:405:0.2) --++ (-3pt,3pt);

		\draw[opacity=0, color=red, ->] (V0) -- (V1);

		\coordinate (CC0) at (1,1.25);
          \draw[opacity=0, color=red, ->] (V0) + (135:0.1) arc (135:405:0.1) --++ (-2pt,2pt);
\end{tikzpicture}