### 2016-04-02 stuff/bspEllipsoidQTensor/bsp.tex
...

\documentclass[a4paper,11pt]{scrartcl}
\usepackage{hyperref}
\usepackage{color}
\input{declarations}

\newcommand{\qspace}{\mathcal{Q}(\surf)}
\newcommand{\tspace}{\mathcal{T}^{(2)}}
\newcommand{\Tr}{\text{Tr}}
\newcommand{\qproject}{\pi_{\qspace}}
\newcommand{\Mb}{\mathbf{M}}

\title{Notes On q-Projection and Helmholtz-deRham-Equation on ellipsoid} 
\author{Ingo Nitschke}

\begin{document}
\maketitle

\section{q-Projection}

We call \(\qspace\) the space of trace-free symmetric surface 2-tensors, shortly surface q-tensors, \ie
\begin{align}
  \qspace &= \left\{ \qb\in\tspace(\surf)\subset \tspace(\R^{3}) \cong \R^{3\times 3}  \middle| \Tr(\qb)=0, \qb^{T}=\qb\right\} \\
          &= \left\{ \qb\in\tspace(\R^{3}) \middle| \Tr(\qb)=0, \qb^{T}=\qb, \qb\cdot\surfNormal=0 (,\surfNormal\cdot \qb=0) \right\}
          \formPeriod
\end{align}
With the surface projection
\begin{align}
  \ProjectSurf = \mathbf{I}-\surfNormal\otimes\surfNormal: \Tangent\R^{3}\cong\R^{3} \rightarrow \Tangent\surf
\end{align}
we introduce the q-projection \( \qproject: \tspace(\R^{3}) \cong \R^{3\times 3} \rightarrow \qspace \) with
\begin{align}
  \qproject(\Mb) := \ProjectSurf\cdot\left( \Mb + \Mb^{T} \right)\cdot\ProjectSurf 
                                - \left( \ProjectSurf : \Mb \right)\ProjectSurf\formPeriod
\end{align}
It is clear, that \( \qproject(\Mb) \) is a surface tensor and symmetric, because \( \ProjectSurf \) is a symmetric surface tensor.
With \( \Tr(\ProjectSurf\cdot\Mb\cdot\ProjectSurf) = \ProjectSurf : \Mb = \Tr(\ProjectSurf\cdot\Mb^{T}\cdot\ProjectSurf) \)
and \( \Tr(\ProjectSurf)=2 \) we obtain the trace-freeness of \( \qproject(\Mb)\).
\ingo{Es waere zu ueberlegen ob man beim q-tensor Modell für \( \R^{3} \)-Tensoren nur die schwaechere(?) Bedingung fordert,
dass der surface-trace \( \ProjectSurf:\qb \) null sei statt der volle \( \R^{3} \)-Trace \( \Tr(\qb) \) 
(vgl. Analogie zu Richtungsfelder: \( \|\ProjectSurf\pb\|_{\R^{3}}=\|\pb\|_{\Tangent\surf} =1 \) vs. \( \|\pb\|_{\R^{3}}=1 \)).}

\section{Notes on Laplace-deRham and other derivative stuff}

\section{Ellipsoid}

  Let be \( \surf \) a ellipsoid with semi-principal axes \( (1,0.5,1.5) \).

\end{document}