### 2016-10-01 stuff/NavierStokesDiskretisierung/tikz/rotrotStaggeredGrid.tex

\begin{tikzpicture}[>=latex, line width=2pt, scale=2.5]
\usetikzlibrary{calc}
\bf
\edef\n{3}

\pgfmathparse{\n-1}
\edef\nmone{\pgfmathresult}

\pgfmathparse{\n-2}
\edef\nmtwo{\pgfmathresult}

%all the gray (dual) stuff
\begin{scope}[gray,  line width=1pt]
%dual vertices
\foreach \x in {0,...,\nmone} {
	\foreach \y in {0,...,\nmone} {
		\coordinate (V) at (\x+0.5,\y+0.5);
		\fill (V) circle(0.5pt);
	}
}

%dual x-edges
\foreach \x in {0,...,\nmtwo} {
	\foreach \y in {0,...,\nmone} {
		\coordinate (V1) at (\x+0.5,\y+0.5);
		\coordinate (V2) at (\x+1.5,\y+0.5); 
		\draw[<-] (V1) -- (V2);
	}
}

%dual y-edges
\foreach \x in {0,...,\nmone} {
	\foreach \y in {0,...,\nmtwo} {
		\coordinate (V1) at (\x+0.5,\y+0.5);
		\coordinate (V2) at (\x+0.5,\y+1.5); 
		\draw[->] (V1) -- (V2);
	}
}

%all remaining (half) dual edges on the boundaries 
\foreach \xy in {0,...,\nmone} {
	\draw[->,densely dotted] (\xy+0.5, 0) --  (\xy+0.5, 0.5); 
	\draw[->,densely dotted] (\xy+0.5, \n-0.5) --  (\xy+0.5, \n);
	\draw[<-,densely dotted] (0,\xy+0.5) -- (0.5,\xy+0.5);
	\draw[<-,densely dotted] (\n-0.5,\xy+0.5) -- (\n,\xy+0.5);
}
\end{scope}

%primal vertices
\foreach \x in {0,...,\n} {
	\foreach \y in {0,...,\n} {
		\coordinate (V) at (\x,\y);
		\fill (V) circle(1pt);
	}
}

%primal x-edges
\foreach \x in {0,...,\nmone} {
	\foreach \y in {0,...,\n} {
		\coordinate (V1) at (\x,\y);
		\coordinate (V2) at (\x+1,\y); 
		\draw[->] (V1) -- (V2);
	}
}

%primal y-edges
\foreach \x in {0,...,\n} {
	\foreach \y in {0,...,\nmone} {
		\coordinate (V1) at (\x,\y);
		\coordinate (V2) at (\x,\y+1); 
		\draw[->] (V1) -- (V2);
	}
}

%primal edge circumcenters
\foreach \xy in {0,...,\n} {
	\foreach \yx in {0,...,\nmone} {
		\coordinate (VX) at (\yx+0.5,\xy);
		\coordinate (VY) at (\xy,\yx+0.5);
		\fill (VX) circle(1pt);
		\fill (VY) circle(1pt);
	}
}

%schemata
\edef\sh{0.6pt}
\draw[blue, yshift=\sh, xshift=\sh, line width=6pt] (0,1.5) circle(1pt) node[above right]{+1} -- (1,1.5) circle(1pt) node[above right]{-2}  -- (2,1.5) circle(1pt)node[above right]{+1};
\draw[blue, yshift=\sh, xshift=\sh, line width=6pt] (0.5,1)circle(1pt) node[above right]{+1} -- (0.5,2) circle(1pt)node[above right]{-1};
\draw[blue, yshift=\sh, xshift=\sh, line width=6pt] (1.5,1)circle(1pt) node[above right]{-1} -- (1.5,2) circle(1pt)node[above right]{+1};

\draw[red, line width=6pt] (1.5,0) circle(1pt) node[above left]{+1} -- (1.5,1) circle(1pt) node[above left]{-2}  -- (1.5,2) circle(1pt)node[above left]{+1};
\draw[red, line width=6pt] (1,1.5) circle(1pt) node[above left]{+1} -- (2,1.5) circle(1pt)node[above left]{-1};
\draw[red, line width=6pt] (1,0.5) circle(1pt) node[above left]{-1} -- (2,0.5) circle(1pt)node[above left]{+1};

%nodes
%\node[below right] at (1,1) {$v_{i,j}$};
%\node[below right] at (2,1) {$v_{i+1,j}$};

\node[below right] at (1.5,1) {$u^x_{i,j}$};
%\node[above right] at (1.5,0) {$e^x_{i,j-1}$};
%\node[above right] at (1.5,2) {$e^x_{i,j+1}$};

\node[below right] at (1,1.5) {$u^y_{i,j}$};
%\node[below right] at (2,1.5) {$e^y_{i+1,j}$};
%\node[below right] at (1,0.5) {$e^y_{i,j-1}$};
%\node[below right] at (2,0.5) {$e^y_{i+1,j-1}$};

\end{tikzpicture}